---
title: "Class10: Holloween Candy"
author: "Yipeng"
format: pdf
editor: visual
---

In today's class we will examine 538 Candy data and see if this helps us gain some mroe feeling for how PCA and other methods work.

```{r}
candy <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/candy-power-ranking/candy-data.csv", row.names = 1)
head(candy)
```

>Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

>Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

>Q. What are these fruity candy?

We can use the ==
```{r}
rownames(candy[candy$fruity==1,])
```

## How often does my favorite candy win

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

I like Haribo Gold Bears, because that's the only sugar I bought before.
```{r}
candy["Haribo Gold Bears",]$winpercent
```

>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat",]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars",]$winpercent
```

```{r}
skimr::skim(candy)
#library("skimr")
#skim(candy)
```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yeap, the `winpercent` column is on a 0:100 scale and lal others appear to be 0:1 scale

>Q7. What do you think a zero and one represent for the candy$chocolate column?

A zero here means the candy is not classified as containing chocolate, and one means do containing chocolate.

>Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
library(ggplot2)
ggplot(candy, aes(winpercent)) +
  geom_histogram(bins=10)
```

>Q9. Is the distribution of winpercent values symmetrical?

No, the winpercent histograms looks right skewed and bimodal.

>Q10. Is the center of the distribution above or below 50%?

below 50% with a mean and median:

```{r}
mean(candy$winpercent)
median(candy$winpercent)
```

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

To answer this question I will need to:
- "subset" (a.k.a. "select", "filter") the candy dataset to just chocolate candy 
- get their winpercent values,
- and then calculate the mean of these. 
- Then do the same for fruity candy and compare

```{r}
# Filter/select/subset to just choclate rows
chocolate.candy <- candy[as.logical(candy$chocolate),]

# Get their winpercent values
chocolate.winpercent <- chocolate.candy$winpercent

# Calculate their mean winpercent value
mean(chocolate.winpercent)
```
Do the same thing to fruity

```{r}
# Do the same thing to fruity
fruity.candy <- candy[as.logical(candy$fruity),]
fruity.winpercent <- fruity.candy$winpercent
mean(fruity.winpercent)
```
So yes, Chocolate has higher average winpercent than fruity candy.

>Q12. Is this difference statistically significant?

```{r}
t.test(chocolate.winpercent, fruity.winpercent)
```

Since the p-value is lower than 0.05, so there is a significant difference, which means people do like chocolate more. 

## Overall Candy Rankings

There is a base R function called `sort()` for, guess what sorting vectors of input.

```{r}
x <- c(5,2,10)
sort(x, decreasing = TRUE)
```

The buddy function to `sort()` that is often more useful is called `ordered`. It returns the "indices" of the input that would result in it being sorted.


```{r}
order(x)
x[order(x)]
```

>Q13. What are the five least liked candy types in this set?

I can order bby `winpercent`

```{r}
ord <- order(candy$winpercent)
head(candy[ord,],5)
```

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
ord <- order(candy$winpercent, decreasing = TRUE)
head(candy[ord,],5)
```

>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```

To make it more colorful

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

>Q17. What is the worst ranked chocolate candy?

Sixlets

>Q18. What is the best ranked fruity candy?

Starburst

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

To deal with overlapping labels I can use the **ggrepel** package

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=2, max.overlaps = 15)
```
So, Reese's Miniatures is the highest wining rank with lowest price.

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
And NiK L Nip is the most expensive and least popular candy.

## Correlation plot

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

A lot of variables are negatively correlated(the dark red color is the -1 correlation). But fruity candy and chocolate are the most anti-correlated.

>Q23. Similarly, what two variables are most positively correlated?

The two candys and most positively correlated are Bar and chocolate.
The two variables are most positively correlated are chocolate and winpercent.

# 6. Principal Component Analysis

The base R function for PCA is called `prcomp()` and we can set "scale=TRUE/FALSE".

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

The main result of PCA - i.e. the new PC plot( projection of candy on our new PC axis) is contained in `pca$x`

```{r}
plot(pca$x[,1:2])

pc <- as.data.frame(pca$x)

ggplot(pc) +
  aes(PC1,PC2, label = rownames(pc)) +
  geom_point(col = my_cols) +
  geom_text_repel(size=2.5, col=my_cols, max.overlaps = 5)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
Fruity, hard and pluribus are picked up strongly by PC1 in positive direction.





































